Minimax-optimal nonparametric regression with graph Laplacians

### 2020-08-09 notes/graph_regression2.tex
Added section on minimax testing with graph regression methods.

\begin{document}
\title{Minimax-optimal learning with graph Laplacians}
\author{Alden Green}
\date{\today}
\maketitle

@@

\textcolor{red}{TODO:} Paragraph motivating graph Laplacian methods.

@@

In graph based learning, we perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$, which should approximately capture the geometry of $P$ and $\mc{X}$, and then constructing an estimate $\wh{f}$ which is smooth with respect to $G_{n,r}$. For a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$. The weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ is an $n \times n$ matrix with entries
\begin{equation*}
\label{eqn:neighborhood_graph}
{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\Biggr),
\end{equation*}
and the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$.  Our star operator, the neighborhood graph Laplacian, can then be written as
\begin{equation}
\label{eqn:graph_Laplacian}
\Lap_{n,r} = \bf{D} - \bf{W}
\end{equation}
\emph{Laplacian eigenmaps} and \emph{Laplacian smoothing} each use $\Lap_{n,r}$ to construct an estimator $\wh{f}$ of $f_0$. Letting $v_1(G_{n,r}),\ldots,v_n(G_{n,r})$ denote eigenvectors of $\Lap_{n,r}$, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ \citep{belkin2003} is given by 
\begin{equation}
\label{eqn:laplacian_eigenmaps}
\wh{f}_{\textrm{LE}} = \sum_{k = 1}^{\kappa} \Dotp{Y}{v_k(G_{n,r})}_n v_{k}(G_{n,r})
\end{equation}
where $\kappa \in \{1,...,n\}$ is a tuning parameter. (Here $\dotp{u}{v}_n = \frac{1}{n}\sum_{i = 1}^{n} u_i v_i$ for vectors $u,v \in \Reals^n$.) The Laplacian smoothing estimator $\wt{f}_{\LS}$ \citep{smola2003} is given by
\begin{equation}
\label{eqn:laplacian_smoothing}
\wt{f}_{\LS}= \min_{f \in \Reals^n} \Bigl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T L^s f \Bigr\}
\end{equation}
where $\rho > 0$ acts a tuning parameter. Then, assuming~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are reasonable estimators of $f_0$, the size of the statistics
\begin{align}
T_{\LE} & = \norm{\wh{f}_{\LE}}_n^2 \label{eqn:laplacian_eigenmaps_test}\\
T_{\LS} & = \norm{\wt{f}_{\LS}}_n^2 \label{eqn:laplacian_smoothing_test}
\end{align}
can be used to test whether or not $f_0 = 0$.

\begin{itemize}
	\item \textbf{Motivate graph Laplacians.} There already exist a vast number of methods for both the nonparametric testing and, especially, the nonparametric estimation problems. Nevertheless, methods such as~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} deserve a place in the nonparametric regression toolbox. Their advantages include:
	\begin{itemize}
		\item Simplicity.
		\item Computational efficiency.
		\item Domain and density adaptivity.
		\item Practical success.
	\end{itemize}
	\item \textbf{Our contributions, short version.} For these reasons, there is a  substantial body of work assessing the statistical properties of graph Laplacian based methods. \textcolor{red}{(TODO): Include citations.} In this paper, we adopt a different perspective, and analyze~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} through the minimax lens. We show that for a variety of function classes $\mc{F}$, distributions $P$, and choices of domain $\mc{X}$ for which minimax rates are known, both~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are minimax optimal estimators (with loss measured in a suitable norm) and both~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} are minimax optimal test statistics. This lends additional theoretical support to the case for using graph Laplacian methods.
\end{itemize}

@@

Before we can state our main results, we must formally define minimaxity in both the estimation and testing problems. As we do so, we will remind the reader of the minimax optimal rates for some classical function classes $\mc{F}$ that we will henceforth consider. We will also briefly recall some (of the many) estimators and tests which achieve these rates; in particular we will focus on the \emph{truncated series} and \emph{smoothing spline} methods, which can be viewed as analogous to the graph Laplacian methods we consider.

\begin{itemize}
	\item \textbf{Minimax estimation.} For given integers $s > 0$ and $p > 0$, the Sobolev space $W^{s,p}(P)$ consists of all functions $f \in \Leb^p(P)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d)$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^p(P)$. The Sobolev $\{s,p\}$ norm is then 
	\begin{equation*}
	\norm{f}_{W^{s,p}(P)}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{X}} \abs{D^{\alpha}f}^p \,dP
	\end{equation*}
	and for a given $M > 0$, the corresponding ball is $W^{s,p}(P, M) = \set{f: \norm{f}_{W^{s,p}(P)} \leq M}$. In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space; we adopt the usual convention of writing $H^s(P) = W^{s,p}(P)$ and $H^s(P,M) = W^{s,2}(P,M)$. We will confine our attention to this special case hereafter. Let us also consider the following assumption

	\begin{enumerate}[(P1)]
		\item 
		\label{asmp:bounded_density}
		The distribution $P$ is defined on domain $\Xset = [0,1]^d$, and admits a density $p$ with respect to the uniform (Lebesgue) measure on $[0,1]^d$. The density $p$ is bounded away from $0$ and $\infty$,
		\begin{equation*}
		0 < p_{\min} < p(x) < p_{\max} < \infty
		\end{equation*}
		for all $x \in [0,1]^d$.
	\end{enumerate}
	The assumption~\ref{asmp:bounded_density} is a standard but significant assumption; we will postpone a more significant discussion of it until \textcolor{red}{a later point}. Assuming~\ref{asmp:bounded_density}, the minimax estimation rate over the Sobolev ellipsoids $H^{s}(P,M)$ is
	\begin{equation*}
	\inf_{\wh{f}} \sup_{f \in H^s(P,M)} \Ebb\Bigl[\norm{\wh{f} - f}_{L^2(P)}^2\Bigr] \asymp n^{-2s/(2s + d)}~~\textrm{for all $s$ and $d$,}
	\end{equation*}
	where the infimum is over all measurable maps $\wh{f}(X_1,Y_1,\ldots,X_n,Y_n)$
	\item \textbf{Minimax hypothesis testing.}
	\item \textbf{Our contributions, long version.}
	\begin{itemize}
		\item \textbf{Minimax optimal estimation.}
		We show that with high probability,  $\norm{\wh{f}_{\LE} - f_0}_{n}^2 \lesssim n^{-2s/(2s + d)}$ whenever $f_0 \in C^s(\mc{X})$ for any $s \leq 1$, or $f_0 \in H^s(P)$ for $s = 1$. The same conclusions hold with respect to $\wt{f}_{\LS}$ when $d \leq 4$. 
		\item \textbf{Minimax optimal testing.}
		We show that with high probability, a level-$\alpha$ test constructed using $T_{\LE}$ has non-trivial power whenever $f_0 \in C^s(\mc{X})$ for any $s \leq 1$---or $f_0 \in H^s(P)$ for $s = 1$---additionally satisfies $\norm{f_0}_{\Leb^2(P)} \gtrsim \max\{n^{-4s/(4s + d)},n^{-1/4}\}$. \textcolor{red}{The same conclusions hold with respect to $\wh{f}_{LS}$.}
		\item \textbf{Manifold adaptivity.}
		If $\mc{X} \subset \Rd$ is a submanifold of dimension $m < d$, each of the aforementioned rates hold with $d$ replaced by $m$.
		\item \textbf{Taking advantage of higher-order smoothness conditions.} Under appropriate boundary conditions on the function classes $C^s$ and $H^s$ and additional smoothness conditions on $P$, when $s > 1$  each of the aforementioned conclusions hold for certain values of $d$.
	\end{itemize}
	\item \textbf{Organization.}
	\item \textbf{Notation.}
\end{itemize}

\section{Minimax optimal graph Laplacian methods}

Let us start by assuming $f_0 \in H^1(P,M)$. In this case, the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and the resulting test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test}, achieve the standard minimax optimal estimation and hypothesis testing rates, respectively.

@@

Under appropriate conditions, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ achieves minimax rates over all $f_0 \in H^1(P,M)$ with high probability. The conditions we impose will be on the neighborhood graph $G_{n,r}$. As the construction of the graph is in the user's control, these assumptions are therefore mild.
\begin{enumerate}[label=(K\arabic*)]
	\item 
	\label{asmp:kernel_form}
	The kernel function $K(z) = \1\Bigl\{\norm{z}_2/r \leq 1 \Bigr\}$.
	\item 
	\label{asmp:kernel_radius}
	The radii $r = r(n)$ satisfies $r(n) \to 0$ as $n \to \infty$, and additionally
	\begin{equation*}
	n^{-1/d}\leq r(n) \leq M^{2/(2 + d)}n^{-1/(2 + d)}
	\end{equation*}
	for all $n$ sufficiently large.
\end{enumerate}

We use~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius} to ensure that the approximation error (bias) of $\wh{f}_{\LE}$ is of the correct order. On the other hand, bounding the estimation error of $\wh{f}_{\LE}$ -- both in expectation and in probability-- is straightforward. Together, bounds on these two sources of error imply Theorem~\ref{thm:laplacian_eigenmaps_estimation1}.
\begin{theorem}
	\label{thm:laplacian_eigenmaps_estimation1}
	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}. Suppose that $f_0 \in H^1(P,M)$, that $P$ satisfies~\ref{asmp:bounded_density}, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius}. Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as given by~\eqref{eqn:laplacian_eigenmaps}, when $\kappa = M^{2/(2 + d)}n^{d/(2 + d)}$, the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
	\begin{equation*}
	\Bigl\|\wh{f}_{\LE} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4 + d)} n^{-2/(2 + d)}
	\end{equation*}
	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp(-\kappa)$.
\end{theorem}

\begin{itemize}
	\item Discuss the choice of radius $r$ imposed by~\ref{asmp:kernel_radius}.
	\item Comment on the distinction between error rates as measured by the norm in $L^2(P_n)$ vs. $L^2(P)$. 
\end{itemize}

Theorem~\ref{thm:laplacian_smoothing_estimation1} is analogous to Theorem~\ref{thm:laplacian_eigenmaps_estimation1}, dealing with the Laplacian smoothing estimator $\wt{f}_{\LS}$. The key difference is that it holds only when the dimension $d < 4$.  
\begin{theorem}
	\label{thm:laplacian_smoothing_estimation1}
	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}. Suppose that $f_0 \in H^1(P,M)$, that $P$ satisfies~\ref{asmp:bounded_density}, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius}. Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as given by~\eqref{eqn:laplacian_eigenmaps}, when $\kappa = M^{2/(2 + d)}n^{d/(2 + d)}$, the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
	\begin{equation*}
	\Bigl\|\wh{f}_{\LE} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4 + d)} n^{-2/(2 + d)}
	\end{equation*}
	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp(-\kappa)$.
\end{theorem}

\begin{itemize}
	\item A restriction on the maximum value of dimension $d$ such as $d \leq 4$ is unsurprising. In a related context (estimation over a $d$-dimensional lattice graph), \cite{sadhanala2016} require the same restriction. The underlying phenomenon is similar to that elucidated by the seminal papers~\cite{birge1993} and~\cite{birge1998}: when the dimension $d$ is too large, the entropy function is irregular and (penalized) least squares approaches are no longer minimax optimal.
\end{itemize}

@@

\section{Higher-order smoothness classes.}
\label{sec:higher_order_smoothness_classes}
Suppose we are willing to assume $f_0$ possesses more than one smooth derivative. We can show that in this situation, the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} obtain the faster (known) minimax rates in the nonparametric estimation and testing problems. However, these conclusions hold only for certain values of the dimension $d$.

\subsection{Minimax optimal estimation}
Under  conditions, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ achieves minimax rates over all $f_0 \in H_0^s(P,M)$ with high probability. These conditions will take two forms. The first pair of conditions will be on the neighborhood graph $G_{n,r}$. These assumptions are mild, as the construction of the graph is in the user's control.
\begin{enumerate}[label=(K\arabic*)]
	\item 
	\label{asmp:kernel_form}
	The kernel function $K(z) = \1\{\norm{z}_2/r \leq 1 \}$.
	\item 
	\label{asmp:kernel_radius}
	The radii $r = r(n)$ satisfy $r(n) \to 0$ as $n \to \infty$, and additionally
	\begin{equation*}
	n^{-1/(2(s - 1) + d)}\leq r(n) \leq M^{2/(2s + d)}n^{-1/(2s + d)}
	\end{equation*}
	for all $n$ sufficiently large.
\end{enumerate}

>>>>

\begin{document}
\title{Minimax-optimal nonparametric regression with graph Laplacians}
\author{Alden Green}
\date{\today}
\maketitle

@@

@@

In graph based nonparametric regression, the idea is to perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense, and then constructing an estimate $\wh{f}$ which is smooth with respect to the graph $G_{n,r}$. For a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$. The weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ is an $n \times n$ matrix with entries
\begin{equation*}
\label{eqn:neighborhood_graph}
{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\Biggr),
\end{equation*}
and the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$.  Our star operator, the neighborhood graph Laplacian, can then be written as
\begin{equation}
\label{eqn:graph_Laplacian}
\Lap_{n,r} = \bf{D} - \bf{W}
\end{equation}
The two graph-based regression methods we will consider are \emph{Laplacian eigenmaps} and \emph{Laplacian smoothing}. Each use $\Lap_{n,r}$ to construct an estimator $\wh{f}$ of $f_0$. Letting $v_1(G_{n,r}),\ldots,v_n(G_{n,r})$ denote eigenvectors of $\Lap_{n,r}$, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ \citep{belkin2003} is given by 
\begin{equation}
\label{eqn:laplacian_eigenmaps}
\wh{f}_{\textrm{LE}} = \sum_{k = 1}^{\kappa} \Dotp{Y}{v_k(G_{n,r})}_n v_{k}(G_{n,r})
\end{equation}
where $\kappa \in \{1,...,n\}$ is a tuning parameter. (Here $\dotp{u}{v}_n = \frac{1}{n}\sum_{i = 1}^{n} u_i v_i$ for vectors $u,v \in \Reals^n$.) The Laplacian smoothing estimator $\wt{f}_{\LS}$ \citep{smola2003} is given by
\begin{equation}
\label{eqn:laplacian_smoothing}
\wt{f}_{\LS}= \min_{f \in \Reals^n} \Bigl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T L^s f \Bigr\}
\end{equation}
where $\rho > 0$ acts a tuning parameter. Then, assuming~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are reasonable estimators of $f_0$, the size of the statistics
\begin{equation}
T_{\LE} = \Bigl\|\wh{f}_{\LE}\Bigr\|_n^2 \label{eqn:laplacian_eigenmaps_test}
\end{equation}
and
\begin{equation}
T_{\LS} = \Bigl\|\wt{f}_{\LS}\Bigr\|_n^2 \label{eqn:laplacian_smoothing_test}
\end{equation}
can be used to test whether or not $f_0 = 0$.

\begin{itemize}
	\item \textbf{Motivate graph Laplacians.} There already exist a vast number of methods for both the nonparametric testing and, especially, the nonparametric estimation problems. Nevertheless, methods such as~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} deserve a place in the nonparametric regression toolbox. Their advantages include:
	\begin{itemize}
		\item Simplicity.
		\item Computational efficiency.
		\item Domain and density adaptivity.
		\item Practical success.
	\end{itemize}
	\item \textbf{Our contributions, short version.} For these reasons, there is a  substantial body of work assessing the statistical properties of graph Laplacian based methods. \textcolor{red}{(TODO): Include citations.} In this paper, we adopt a different perspective, and analyze~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} through the minimax lens. We show that for a variety of function classes $\mc{F}$, distributions $P$, and choices of domain $\mc{X}$ for which minimax rates are known, both~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are minimax optimal estimators (with loss measured in a suitable norm) and both~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} are minimax optimal test statistics. This lends additional theoretical support to the case for using graph Laplacian methods.
	\item \textbf{Summary}
\end{itemize}

@@

Before we get to our main results, we will review some of the relevant literature on nonparametric estimation and regression, and recall some relevant minimax analysis.  We will also briefly recall some (of the many) estimators and tests which achieve these rates. In particular, we will focus on the \emph{truncated series} and \emph{smoothing spline} methods. As we will see, Laplacian eigenmaps and Laplacian smoothing can be seen as graph-based analogues to these ``continuuum'' methods.

\begin{itemize}
	\item \textbf{Minimax estimation.} For given integers $s > 0$ and $p > 0$, the Sobolev space $W^{s,p}(P)$ consists of all functions $f \in \Leb^p(P)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d)$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^p(P)$. The Sobolev $\{s,p\}$ norm is then 
	\begin{equation*}
	\norm{f}_{W^{s,p}(P)}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{X}} \abs{D^{\alpha}f}^p \,dP
	\end{equation*}
	and for a given $M > 0$, the corresponding ball is $W^{s,p}(P, M) = \set{f: \norm{f}_{W^{s,p}(P)} \leq M}$. In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space; we adopt the usual convention of writing $H^s(P) = W^{s,p}(P)$ and $H^s(P,M) = W^{s,2}(P,M)$. We will confine our attention to this special case hereafter. Let us also make the following assumption:
	\begin{enumerate}[(P1)]
		\item 
		\label{asmp:bounded_density}
		The distribution $P$ is defined on domain $\Xset = [0,1]^d$, and admits a density $p$ with respect to the uniform (Lebesgue) measure on $[0,1]^d$. The density $p$ is bounded away from $0$ and $\infty$,
		\begin{equation*}
		0 < p_{\min} < p(x) < p_{\max} < \infty
		\end{equation*}
		for all $x \in [0,1]^d$.
	\end{enumerate}
	The assumption~\ref{asmp:bounded_density} is a standard but significant assumption; we will return to discuss it further at \textcolor{red}{a later point}. Assuming~\ref{asmp:bounded_density}, the minimax estimation rate over the Sobolev ellipsoids $H^{s}(P,M)$ is
	\begin{equation*}
	\inf_{\wh{f}} \sup_{f \in H^s(P,M)} \Ebb\Bigl[\norm{\wh{f} - f}_{L^2(P)}^2\Bigr] \asymp n^{-2s/(2s + d)}~~\textrm{for all $s$ and $d \geq 0$,}
	\end{equation*}
	where the infimum is over all measurable maps $\wh{f}(X_1,Y_1,\ldots,X_n,Y_n)$
	\item \textbf{Minimax hypothesis testing.}
	\item \textbf{Our contributions, long version.}
	\begin{itemize}
		\item \textbf{Minimax optimal estimation.}
		We show that with high probability,  $\norm{\wh{f}_{\LE} - f_0}_{n}^2 \lesssim n^{-2s/(2s + d)}$ whenever $f_0 \in C^s(\mc{X})$ for any $s \leq 1$, or $f_0 \in H^s(P)$ for $s = 1$. The same conclusions hold with respect to $\wt{f}_{\LS}$ when $d \leq 4$. 
		\item \textbf{Minimax optimal testing.}
		We show that with high probability, a level-$\alpha$ test constructed using $T_{\LE}$ has non-trivial power whenever $f_0 \in C^s(\mc{X})$ for any $s \leq 1$---or $f_0 \in H^s(P)$ for $s = 1$---additionally satisfies $\norm{f_0}_{\Leb^2(P)} \gtrsim \max\{n^{-4s/(4s + d)},n^{-1/4}\}$. \textcolor{red}{The same conclusions hold with respect to $\wh{f}_{LS}$.}
		\item \textbf{Manifold adaptivity.}
		If $\mc{X} \subset \Rd$ is a submanifold of dimension $m < d$, each of the aforementioned rates hold with $d$ replaced by $m$.
		\item \textbf{Taking advantage of higher-order smoothness conditions.} Under appropriate boundary conditions on the function classes $C^s$ and $H^s$ and additional smoothness conditions on $P$, when $s > 1$  each of the aforementioned conclusions hold for certain values of $d$.
	\end{itemize}
	\item \textbf{Notation.}
\end{itemize}

\section{Minimax optimal graph Laplacian methods}
\label{sec:minimax_optimal_graph_Laplacian_methods}

Let us start by assuming $f_0 \in H^1(P,M)$. In this section, we will establish that the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and the resulting test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test}, achieve the standard minimax optimal estimation and hypothesis testing rates, respectively.

@@

Under appropriate conditions, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ achieves minimax rates over all $f_0 \in H^1(P,M)$ with high probability. The conditions we impose will be on the neighborhood graph $G_{n,r}$. As the construction of the graph is in the user's control, these assumptions are therefore mild.
\begin{enumerate}[label=(K\arabic*)]
	\item 
	\label{asmp:kernel_form}
	The kernel function $K(z) = \1\Bigl\{\norm{z}_2/r \leq 1 \Bigr\}$.
	\item 
	\label{asmp:kernel_radius_estimation}
	The radii $r = r(n)$ satisfies $r(n) \to 0$ as $n \to \infty$, and additionally
	\begin{equation*}
	n^{-1/d}\leq r(n) \leq M^{2/(2 + d)}n^{-1/(2 + d)}
	\end{equation*}
	for all $n$ sufficiently large.
\end{enumerate}

We use~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_estimation} to ensure that the approximation error (bias) of $\wh{f}_{\LE}$ is of the correct order. On the other hand, bounding the estimation error of $\wh{f}_{\LE}$ -- both in expectation and in probability-- is straightforward. Together, bounds on these two sources of error imply Theorem~\ref{thm:laplacian_eigenmaps_estimation1}.
\begin{theorem}
	\label{thm:laplacian_eigenmaps_estimation1}
	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}. Suppose that $f_0 \in H^1(P,M)$, that $P$ satisfies~\ref{asmp:bounded_density}, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_estimation}. Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as given by~\eqref{eqn:laplacian_eigenmaps}, when $\kappa = M^{2/(2 + d)}n^{d/(2 + d)}$, the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
	\begin{equation*}
	\Bigl\|\wh{f}_{\LE} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4 + d)} n^{-2/(2 + d)}
	\end{equation*}
	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp(-\kappa)$.
\end{theorem}

\begin{itemize}
	\item Discuss the choice of radius $r$ imposed by~\ref{asmp:kernel_radius_estimation}.
	\item Comment on the distinction between error rates as measured by the norm in $L^2(P_n)$ vs. $L^2(P)$.
	\item \textbf{Proof techniques}: to prove results we rely on both (a) classical bias-variance decomposition of orthogonal series estimators (see e.g. \textcolor{red}{Tsybakov}), and (b) pointwise and variational results on neighborhood graph Laplacian semi-norms and eigenvalues (see e.g. \textcolor{red}{Belkin04, Belkin08, GarciaTrillos18, Calder19})
\end{itemize}

Theorem~\ref{thm:laplacian_smoothing_estimation1} is analogous to Theorem~\ref{thm:laplacian_eigenmaps_estimation1}, except it deals with the Laplacian smoothing estimator $\wt{f}_{\LS}$ rather than the Laplacian eigenmaps estimator $\wh{f}_{\LE}$. The key difference is that it holds only when the dimension $d < 4$.  
\begin{theorem}
	\label{thm:laplacian_smoothing_estimation1}
	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}. Suppose that $f_0 \in H^1(P,M)$, that $P$ satisfies~\ref{asmp:bounded_density}, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_estimation}. Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as given by~\eqref{eqn:laplacian_eigenmaps}, when we set $\kappa = M^{2/(2 + d)}n^{d/(2 + d)}$ the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
	\begin{equation*}
	\Bigl\|\wh{f}_{\LS} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4 + d)} n^{-2/(2 + d)}
	\end{equation*}
	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp(-\kappa)$.
\end{theorem}

\begin{itemize}
	\item \textbf{Proof techniques}: Because the solution to~\eqref{eqn:laplacian_smoothing} has an analytic form, we can rely on a bias-variance decomposition---as opposed to empirical process theory---to derive~Theorem~\ref{thm:laplacian_smoothing_estimation1}.
	\item The restriction in Theorem~\ref{thm:laplacian_smoothing_estimation1} that the dimension $d$ be less than $4$ is unsurprising. In a related context (estimation over a $d$-dimensional lattice graph), \cite{sadhanala2016} require the same restriction. The underlying phenomenon is similar to that elucidated by the seminal papers~\cite{birge1993} and~\cite{birge1998}: when the dimension $d$ is at least $4$, the parameter space is too large, and (penalized) least squares approaches are no longer minimax optimal.
\end{itemize}

@@

Under appropriate conditions, the Laplacian eigenmaps test statistic $T_{\LE}$ has non-trivial power over all $f_0 \in H^1(P,M)$ that are separated from $0$ in $\Leb^2(P)$ norm by at least the minimax critical radius. These conditions include~\ref{asmp:kernel_form} and~\ref{asmp:smooth_density}, but not~\ref{asmp:kernel_radius_estimation}. Instead, we will require the following scaling of the neighborhood graph radius.
\begin{enumerate}[label=(K\arabic*)]
	\setcounter{enumi}{2}
	\item 
	\label{asmp:kernel_radius_testing}
	The radii $r = r(n)$ satisfy $r(n) \to 0$ as $n \to \infty$, and additionally
	\begin{equation*}
	n^{-1/d} \leq r(n) \leq n^{-2/(4 + d)}
	\end{equation*}
	for all $n$ sufficiently large.
\end{enumerate}

\begin{theorem}
	\label{thm:laplacian_eigenmaps_testing1}
	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}. Suppose that $f_0 \in H^1(P,M)$, that
	$P$ satisfies~\ref{asmp:bounded_density}, and that the neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_testing}.
	Then for the Laplacian eigenmaps test statistic $T_{\LE}$ as in~\eqref{eqn:laplacian_eigenmaps_test} with $\kappa = n^{2d/(4 + d)} M^{4d/(4 + d)}$, the following statement holds true for all $n$ sufficiently large: there exist constants $C_1$ and $C_2$ which do not depend on $f_0$ such that if
	\begin{equation}
	\label{eqn:laplacian_eigenmaps_testing}
	\norm{f_0}_{\Leb^2(P)}^2 \geq \frac{C_1}{\delta^2} M^{2d/(4 + d)} n^{-4/(4 + d)}
	\end{equation}
	then
	\begin{equation*}
	\Pbb \biggl(T_{\LE} \geq \frac{\kappa}{n} + \frac{1}{\delta}\sqrt{\frac{2\kappa}{n^2}}\biggr) \geq 1 - C_2\Bigl(\delta + r^{-d} \exp\{- c n r^d\}\Bigr).
	\end{equation*}
\end{theorem}
\begin{itemize}
	\item Explain why Theorem~\ref{thm:laplacian_eigenmaps_testing1} implies minimax optimality.
\end{itemize}

The estimation and testing rates derived in Section~\ref{sec:minimax_optimal_graph_Laplacian_methods} exhibit the curse of dimensionality: as $d$ gets large, inference becomes hard. As covered in Section~\ref{sec:problem_setup_and_background}, it is well known that by making additional structural assumptions---either on the regression function $f_0$, the distribution $P$, or both---one can derive improved minimax rates. In the following two sections, we show that under such assumptions, the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} can in certain situations obtain the faster (known) minimax rates in the nonparametric estimation and testing problems.

\section{Higher-order smoothness classes}
\label{sec:higher_order_smoothness_classes}
In this section, we will assume that $f_0 \in H^s(P,M)$ for some integer $s > 1$. In other words, we will assume that $f_0$ has at least two smooth derivatives.

\subsection{Minimax optimal estimation}
Under the following three conditions, the estimation error of the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ achieves the minimax optimal rate for any $f_0 \in H^s(P,M)$.
\begin{enumerate}[label=(K\arabic*)]
	\setcounter{enumi}{2}
	\item 
	\label{asmp:kernel_radius_higher_order}
	\textbf{Neighborhood graph radius}:
	The sequence of radii $r = r(n)$ satisfy $r(n) \to 0$ as $n \to \infty$, and additionally
	\begin{equation*}
	n^{-1/(2(s - 1) + d)}\leq r(n) \leq M^{2/(2s + d)}n^{-1/(2s + d)}
	\end{equation*}
	for all $n$ sufficiently large.
\end{enumerate}
\begin{enumerate}[label=(F\arabic*)]
	\setcounter{enumi}{0}
	\item 
	\label{asmp:boundary} \textbf{Boundary condition}; The regression function $f_0$ is zero-trace. In other words, $f_0 \in H_0^s(P,M)$. 
\end{enumerate}
\begin{enumerate}[label=(P\arabic*)]
	\setcounter{enumi}{1}
	\item 
	\label{asmp:smooth_density}\textbf{$\beta$-Holder density}: The design distribution $P$ has a density $p \in C^{\beta}([0,1]^d;p_{\max})$ for some $\beta \geq 0$. 
\end{enumerate}
